---
title: "CUNY DATA 621 - Business Analytics and Data Mining"
subtitle: "Week 1 - Recommended"
author: "Walt Wells, 2018"
output:
  html_document:
    css: ../custom.css
    highlight: zenburn
    theme: lumen
---

# Environment Prep

```{r, message=F, warning=F}
if (!require('faraway')) install.packages('faraway')
```


# 1.1

The dataset teengamb concerns a study of teenage gambling in Britain.  Make a numerical and graphical summary of the data, commenting on any features that you find interesting. Limit the output you present to a quantity that a busy reader would find sufficient to get a basic understanding of the data.

```{r}
teengamb$sex <- factor(teengamb$sex)
levels(teengamb$sex) <- c('male', 'female')
summary(teengamb)
plot(teengamb[,2:5])
plot(gamble~sex, teengamb)
```

# 1.3

The dataset prostate is from a study on 97 men with prostate cancer who were due to receive a radical prostatectomy. Make a numerical and graphical summary of the data as in the first question.

```{r}
prostate$svi <- factor(prostate$svi)
prostate$gleason <- factor(prostate$gleason)
summary(prostate)
plot(prostate)
```

# 1.4

The dataset sat comes from a study entitled “Getting What You Pay For: The Debate Over Equity in Public School Expenditures.”  Make a numerical and graphical summary of the data as in the first question.

```{r}
summary(sat)
plot(sat)
```

# 1.5 
The dataset divusa contains data on divorces in the United States from 1920 to 1996. Make a numerical and graphical summary of the data as in the first question.

```{r}
divusa$year <- factor(divusa$year)
summary(divusa)
plot(divusa)
```